# Results

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(ggplot2)
library(tidyverse)
library(stringr)
library(GGally)
library(parcoords)
library(tidyr)
library(r2d3)
library(dplyr)
library(ggmap)
library(maps)
library(mapdata)
library(readr)
library(gridExtra)
```

```{r include=FALSE}
# read file
gun_vio=read.csv("../dataset/clean gun-violence-data_2013-3-2018-3.csv")
gun_vio["year"]=str_sub( as.character(gun_vio$date),1,4)
gun_vio<-gun_vio[which(as.numeric(as.character(gun_vio$year))>2013)  ,]
gun_vio$date<-as.Date(gun_vio$date)
gun_vio$state<-as.character(gun_vio$state)
gun_vio["year"]<-format(as.Date(gun_vio$date, format="%d/%m/%Y"),"%Y")
gun_vio["month"]<-format(as.Date(gun_vio$date, format="%d/%m/%Y"),"%m")

gun_vio_n <- read.csv("../dataset/gun_violence_sum.csv")
back_check=read.csv("../dataset/nics-firearm-background-checks copy.csv")
GDP<-read.csv("../dataset/GDP_states.csv")
GDP_sub<-GDP %>% select(-c(1,2,4,5,6,7,8,9))

gun_character <- read.csv("../dataset/gun-violence-character.csv")
#uem_month=read.csv("dataset/unemployment_states_by_month.csv")
population_tp=read.csv("../dataset/clean_population_by_state.csv")
population<-select(population_tp,2,3,4)

back_check$month<-as.character(back_check$month)
back_check["year"]=str_sub(back_check$month,1,4)


```

## Exploring the Gun Violence Data

We start from an interesting question: Where is the least gun threaded state in the US? Then we would like explore the gun violence data, hoping to find the answers from sub questions. 

### How many gun violence incidents happened every year in each state?

We calculate the the sum of of incidents number of each state from 2014 to 2018, the plots are shown below, 
```{r include=FALSE}
#incident number
gun_vio_states<-gun_vio%>%
  group_by(state,year)%>%
  summarize(incident_number=n())
gun_vio_states<-as.data.frame(gun_vio_states)
gun_vio_states_2014 <- gun_vio_states %>%
  filter(year=="2014")
gun_vio_states_2015 <- gun_vio_states %>%
  filter(year=="2015")
gun_vio_states_2016 <- gun_vio_states %>%
  filter(year=="2016")
gun_vio_states_2017 <- gun_vio_states %>%
  filter(year=="2017")
gun_vio_states_2018 <- gun_vio_states %>%
  filter(year=="2018")

sum_2014 <- ggplot(gun_vio_states_2014 ,aes(x=fct_reorder(state,incident_number),y=incident_number))+
  geom_bar(position = "dodge", stat = "identity")+
  ylim(0,5100)+
  coord_flip()+
  ggtitle("Incident Number by States in 2015")+
  labs(y="incident number",x="States")

sum_2015 <- ggplot(gun_vio_states_2015 ,aes(x=fct_reorder(state,incident_number),y=incident_number))+
  geom_bar(position = "dodge", stat = "identity")+
  ylim(0,5100)+
  coord_flip()+
  ggtitle("Incident Number by States in 2015")+
  labs(y="incident number",x="States")

sum_2016 <- ggplot(gun_vio_states_2016 ,aes(x=fct_reorder(state,incident_number),y=incident_number))+
  geom_bar(position = "dodge", stat = "identity")+
  ylim(0,5100)+
  coord_flip()+
  ggtitle("Incident Number by States in 2016")+
  labs(y="incident number",x="States")

sum_2017 <- ggplot(gun_vio_states_2017 ,aes(x=fct_reorder(state,incident_number),y=incident_number))+
  geom_bar(position = "dodge", stat = "identity")+
  ylim(0,5100)+
  coord_flip()+
  ggtitle("Incident Number by States in 2017")+
  labs(y="incident number",x="States")

sum_2018 <- ggplot(gun_vio_states_2018 ,aes(x=fct_reorder(state,incident_number),y=incident_number))+
  geom_bar(position = "dodge", stat = "identity")+
  ylim(0,5100)+
  coord_flip()+
  ggtitle("Incident Number by States in 2018")+
  labs(y="incident number",x="States")
```

```{r,fig.height=15,fig.width=9}
grid.arrange(sum_2014,sum_2015,sum_2016,sum_2017,sum_2018,ncol=2)
```

The bar in 2018 is shorter than the bars in other years, it's because we only have the record from Jan to March in 2018. We could find that in 2014 and 2018, California has the most incident number; while in 2015, 2016 and 2017, Illinois has the most incident number. But, does the large number comes from the large population? Maybe it's better to calculate the incident number per 100000 person. The plots are shown below, 

```{r include=FALSE}
#incident number per 100000 person
gun_vio_adjust = merge(gun_vio_states,population, by.x=c("state", "year"), by.y=c("state", "year"))
gun_vio_adjust["incident_num_avg"]<-gun_vio_adjust$incident_number/gun_vio_adjust$population*100000
gun_vio_adjust<-select(gun_vio_adjust,1,2,5)

gun_vio_adjust_2014 <- gun_vio_adjust %>%
  filter(year=="2014")
gun_vio_adjust_2015 <- gun_vio_adjust %>%
  filter(year=="2015")
gun_vio_adjust_2016 <- gun_vio_adjust %>%
  filter(year=="2016")
gun_vio_adjust_2017 <- gun_vio_adjust %>%
  filter(year=="2017")
gun_vio_adjust_2018 <- gun_vio_adjust %>%
  filter(year=="2018")

avg_2014 <- ggplot(gun_vio_adjust_2014 ,aes(x=fct_reorder(state,incident_num_avg),y=incident_num_avg))+
  geom_bar(position = "dodge", stat = "identity")+
  ylim(0,170)+
  coord_flip()+
  ggtitle("Incident Number per 100000 person by States in 2014")+
  labs(y="average incident number",x="States")

avg_2015 <- ggplot(gun_vio_adjust_2015 ,aes(x=fct_reorder(state,incident_num_avg),y=incident_num_avg))+
  geom_bar(position = "dodge", stat = "identity")+
  ylim(0,170)+
  coord_flip()+
  ggtitle("Incident Number per 100000 person by States in 2015")+
  labs(y="average incident number",x="States")

avg_2016 <- ggplot(gun_vio_adjust_2016 ,aes(x=fct_reorder(state,incident_num_avg),y=incident_num_avg))+
  geom_bar(position = "dodge", stat = "identity")+
  ylim(0,170)+
  coord_flip()+
  ggtitle("Incident Number per 100000 person by States in 2016")+
  labs(y="average incident number",x="States")

avg_2017 <- ggplot(gun_vio_adjust_2017 ,aes(x=fct_reorder(state,incident_num_avg),y=incident_num_avg))+
  geom_bar(position = "dodge", stat = "identity")+
  ylim(0,170)+
  coord_flip()+
  ggtitle("Incident Number per 100000 person by States in 2017")+
  labs(y="average incident number",x="States")

avg_2018 <- ggplot(gun_vio_adjust_2018 ,aes(x=fct_reorder(state,incident_num_avg),y=incident_num_avg))+
  geom_bar(position = "dodge", stat = "identity")+
  ylim(0,170)+
  coord_flip()+
  ggtitle("Incident Number per 100000 person by States in 2018")+
  labs(y="average incident number",x="States")
```

```{r,fig.height=15,fig.width=9}
grid.arrange(avg_2014,avg_2015,avg_2016,avg_2017,avg_2018,ncol=2)
```

The resultes turns out very different, District of Columbia moves to the top of the bar from 2014 to 2018; Alaska becomes the second most in 2015 and 2016. It might because the population number in these states are quite small. 

### How many people get injured?

The gun violence may includes mass shooting. In every incident, the number of people who are injured or killed are different, we want to explore this number. The plot is shown below, 

```{r include=FALSE}

gun_vio_n_2014 <- gun_vio_n %>%filter(year=="2014")
gun_vio_n_2015 <- gun_vio_n %>%filter(year=="2015")
gun_vio_n_2016 <- gun_vio_n %>%filter(year=="2016")
gun_vio_n_2017 <- gun_vio_n %>%filter(year=="2017")
gun_vio_n_2018 <- gun_vio_n %>%filter(year=="2018")

sum_n_2014 <- ggplot(gun_vio_n_2014 ,aes(x=fct_reorder(state,total_TOLL),y=total_TOLL))+
  geom_bar(position = "dodge", stat = "identity")+
  ylim(0,5100)+
  coord_flip()+
  ggtitle("Injured Number by States in 2014")+
  labs(y="incident number",x="States")

sum_n_2015 <- ggplot(gun_vio_n_2015 ,aes(x=fct_reorder(state,total_TOLL),y=total_TOLL))+
  geom_bar(position = "dodge", stat = "identity")+
  ylim(0,5100)+
  coord_flip()+
  ggtitle("Injured Number by States in 2015")+
  labs(y="incident number",x="States")

sum_n_2016 <- ggplot(gun_vio_n_2016 ,aes(x=fct_reorder(state,total_TOLL),y=total_TOLL))+
  geom_bar(position = "dodge", stat = "identity")+
  ylim(0,5100)+
  coord_flip()+
  ggtitle("Injured Number by States in 2016")+
  labs(y="incident number",x="States")

sum_n_2017 <- ggplot(gun_vio_n_2017 ,aes(x=fct_reorder(state,total_TOLL),y=total_TOLL))+
  geom_bar(position = "dodge", stat = "identity")+
  ylim(0,5100)+
  coord_flip()+
  ggtitle("Injured Number by States in 2017")+
  labs(y="incident number",x="States")

sum_n_2018 <- ggplot(gun_vio_n_2018 ,aes(x=fct_reorder(state,total_TOLL),y=total_TOLL))+
  geom_bar(position = "dodge", stat = "identity")+
  ylim(0,5100)+
  coord_flip()+
  ggtitle("Injured Number by States in 2018")+
  labs(y="incident number",x="States")
```

```{r,fig.height=15,fig.width=9}
grid.arrange(sum_n_2014,sum_n_2015,sum_n_2016,sum_n_2017,sum_n_2018,ncol=2)
```

The bar in 2018 is shorter than the bars in other years, it's because we only have the record from Jan to March in 2018. Then we would like to look at the this number per 100000 person. The plots are shown below, 

```{r include=FALSE}
gun_vio_n_adjust = merge(gun_vio_n,population, by.x=c("state", "year"), by.y=c("state", "year"))

gun_vio_n_adjust['sum_killed_avg']<-gun_vio_n_adjust$total_TOLL/gun_vio_n_adjust$population*100000
gun_vio_n_adjust<-select(gun_vio_n_adjust,1,2,6)

gun_vio_adjust_2014 <- gun_vio_n_adjust %>%
  filter(year=="2014")
gun_vio_adjust_2015 <- gun_vio_n_adjust %>%
  filter(year=="2015")
gun_vio_adjust_2016 <- gun_vio_n_adjust %>%
  filter(year=="2016")
gun_vio_adjust_2017 <- gun_vio_n_adjust %>%
  filter(year=="2017")
gun_vio_adjust_2018 <- gun_vio_n_adjust %>%
  filter(year=="2018")


avg_n_2014 <- ggplot(gun_vio_adjust_2014 ,aes(x=fct_reorder(state,sum_killed_avg),y=sum_killed_avg))+
  geom_bar(position = "dodge", stat = "identity")+
  ylim(0,70)+
  coord_flip()+
  ggtitle("Injured Number per 100000 person by States in 2014")+
  labs(y="average incident number",x="States")

avg_n_2015 <- ggplot(gun_vio_adjust_2015 ,aes(x=fct_reorder(state,sum_killed_avg),y=sum_killed_avg))+
  geom_bar(position = "dodge", stat = "identity")+
  ylim(0,70)+
  coord_flip()+
  ggtitle("Injured Number per 100000 person by States in 2015")+
  labs(y="average incident number",x="States")

avg_n_2016 <- ggplot(gun_vio_adjust_2016 ,aes(x=fct_reorder(state,sum_killed_avg),y=sum_killed_avg))+
  geom_bar(position = "dodge", stat = "identity")+
  ylim(0,70)+
  coord_flip()+
  ggtitle("Injured Number per 100000 person by States in 2016")+
  labs(y="average incident number",x="States")

avg_n_2017 <- ggplot(gun_vio_adjust_2017 ,aes(x=fct_reorder(state,sum_killed_avg),y=sum_killed_avg))+
  geom_bar(position = "dodge", stat = "identity")+
  ylim(0,70)+
  coord_flip()+
  ggtitle("Injured Number per 100000 person by States in 2017")+
  labs(y="average incident number",x="States")

avg_n_2018 <- ggplot(gun_vio_adjust_2018 ,aes(x=fct_reorder(state,sum_killed_avg),y=sum_killed_avg))+
  geom_bar(position = "dodge", stat = "identity")+
  ylim(0,70)+
  coord_flip()+
  ggtitle("Injured Number per 100000 person by States in 2018")+
  labs(y="average incident number",x="States")
```

```{r,fig.height=15,fig.width=9}
grid.arrange(avg_n_2014,avg_n_2015,avg_n_2016,avg_n_2017,avg_n_2018,ncol=2)
```

### Where does these gun violence happen?

Then, we zoomed the map to find the exact location of each gun incident. We ploted each event as a red dot on the map. The following plot shows the gun violence in California in 2013, 

```{r include=FALSE}
data <- read_csv("../dataset/gun-violence-data_01-2013_03-2018.csv")
names(data)[15] <- "lat"
names(data)[17] <- "lon"
data$date <- as.character(data$date)
data_new <- data %>%
  filter(state=="California")
data_new["year"]=str_sub( as.character(data_new$date),1,4)
data_new <- data_new %>%
  filter(year=="2013")
```

```{r}
qmplot(lon,lat,data=data_new,colour=I("red"),zoom=9,size = I(0.9),alpha=.3,maptype = "toner-lite",main = "Gun violence in California in 2013")+
  theme(legend.position="none")
```

As we could see from this plot, the gun violence incident is clustered at cities and roads, which means that gun violences is mostly happen where people are. 

### Which kind of gun violence is the most common?

We look at the key characteristics of gun violence incidents, sum the same kind of incidents from 2014 to 2018 in every state, the plot is shown below, 

```{r include=FALSE}
character<-gun_character
character['Mass Shooting']<-str_detect(character$incident_characteristics,'Mass Shooting')
character['Shot - Wounded/Injured']<-str_detect(character$incident_characteristics,'Shot - Wounded/Injured')
character['Shot - Dead (murder, accidental, suicide)']<-str_detect(character$incident_characteristics,'Shot - Dead')
character['Possession (gun(s) found during commission of other crimes)']<-str_detect(character$incident_characteristics,'found during commission of other crimes')
character['Possession of gun by felon or prohibited person']<-str_detect(character$incident_characteristics,'Possession of gun by felon or prohibited person')
character['Drug involvement']<-str_detect(character$incident_characteristics,'Drug involvement')
character['Shots Fired']<-str_detect(character$incident_characteristics,'Shots Fired')
character['Bar/club incident - in or around establishment']<-str_detect(character$incident_characteristics,'Bar/club incident - in or around establishment')
character['Armed robbery with injury/death and/or evidence of DGU found']<-str_detect(character$incident_characteristics,'Armed robbery with injury/death and/or evidence of DGU found')
character['Shot - Dead (murder, accidental, suicide)']<-str_detect(character$incident_characteristics,'Dead')
character['Officer Involved Shooting']<-str_detect(character$incident_characteristics,'Officer Involved Shooting')
character['Kidnapping/abductions/hostage']<-str_detect(character$incident_characteristics,'Kidnapping/abductions/hostage')
character['Brandishing/flourishing/open carry/lost/found']<-str_detect(character$incident_characteristics,'Brandishing/flourishing/open carry/lost/found')
character['Murder/Suicide']<-str_detect(character$incident_characteristics,'Murder/Suicide')
character['Drug involvement']<-str_detect(character$incident_characteristics,'Drug involvement')
character['Drive-by (car to street, car to car)']<-str_detect(character$incident_characteristics,'(car to street, car to car)')
character['Defensive Use']<-str_detect(character$incident_characteristics,'Defensive Use')
character['Domestic Violence']<-str_detect(character$incident_characteristics,'Domestic Violence')
character['Home Invasion']<-str_detect(character$incident_characteristics,'Home Invasion')
character['Institution/Group/Business']<-str_detect(character$incident_characteristics,'Institution/Group/Business')
character['Concealed Carry License - Perpetrator']<-str_detect(character$incident_characteristics,'Concealed Carry License - Perpetrator')
character['School Shooting']<-str_detect(character$incident_characteristics,'School Shooting')
character['Car-jacking']<-str_detect(character$incident_characteristics,'Car-jacking')
character['Hate crime']<-str_detect(character$incident_characteristics,'Hate crime')
character['House party']<-str_detect(character$incident_characteristics,'House party')
character['Workplace shooting']<-str_detect(character$incident_characteristics,'Workplace shooting')
character['Gun(s) stolen from owner']<-str_detect(character$incident_characteristics,'Gun(s) stolen from owner')
character['Accidental/Negligent Discharge']<-str_detect(character$incident_characteristics,'Accidental/Negligent Discharge')
character['Accidental Shooting']<-str_detect(character$incident_characteristics,'Accidental Shooting')
character['ATF/LE Confiscation/Raid/Arrest']<-str_detect(character$incident_characteristics,'ATF/LE Confiscation/Raid/Arrest')
character['Non-Shooting Incident']<-str_detect(character$incident_characteristics,'Non-Shooting Incident')
character['Shots Fired - No Injuries']<-str_detect(character$incident_characteristics,'Shots Fired - No Injuries')
character=select(character,-1)
character_sum<-colSums(character[,-1])
character_sum_df_tp<-as.data.frame(as.list(character_sum))
character_sum_df<-gather(character_sum_df_tp)
```

```{r,fig.width=6}
character_sum_df%>%filter(value>1000)%>%
ggplot(aes(x= fct_reorder(key, value)   ,y=value))+
  geom_bar(position = "dodge", stat = "identity")+
  coord_flip()+
  ggtitle("Key characteristics of Gun Violence Incidents")
```

### Is the violence rate stable?

## Effects of Gun Possesion

Since the numbers of firearm possesions are different among states, it might have something to do with the gun violence. 

### How does the gun change rate influence the gun violence status?

## Effects of GDP data

We also would like to find whether there is something to do with the ecnomic situation. We used GDP of each state to represent the economic situation. 